Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2017-08-19 almo-langevin.tex
delta f is introduces, questions for first figure

% Re-write the following section and include any recent advances
In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian, known as canonical KS orbitals, must be replaced with an alternative set of \emph{local} electronic descriptors
~\footnote{Canonical KS orbitals are delocalized over all atoms in the system. 
This delocalization is the main reason for the steep growth of the cost of the conventional KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom that describe delocalized orbitals grows quadratically.}. 
Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM). 
%However, the variational optimization of the DM is very inefficient for accurate DFT calculations that require many basis functions per atom~\cite{a:ls-rev-1999,a:ls-dm-sign}. 
However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM formulation is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
%However, in accurate DFT calculations with many basis functions per atom are required, the DM becomes sufficiently sparse only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}. 
Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis. 
Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix.

Alternatively, LS can be achieved using a set of local electronic descriptors constructed by enforcing strict locality on the KS orbitals and, at the same time, by relaxing the orthogonality constraints imposed on them. 
From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are more promising since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the progress in the development of orbital-based LS methods has been hindered by the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.

%Thus, despite remarkable recent progress in linear scaling density function theory, the computational cost of existing methods remains too high for routine \emph{ab initio} molecular dynamics (AIMD) simulations.
Thus, despite remarkable recent progress the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets. Otherwise, LS DFT cannot compete with the straightforward cubically scaling KS DFT on the length scales accessible in routine AIMD.
%Thus,  the computational cost of the existing asymptotically linear-scaling DFT methods remains too high to compete with the straightforward cubically scaling KS DFT on the length scales practically accessible in routine AIMD simulations.
%Thus, although there exist numerous DFT methods capable of yielding asymptotic LS, their computational cost for systems in/on the practically accessible length scale remains too high to compete with the conventional cubically scaling KS DFT with its highly-optimized numerical algorithms.

In this work, we introduce a new AIMD method for weakly-interacting molecular systems that can reach the linear scaling regime for systems of medium size while maintaining a low computational overhead. Using liquid water as an example, we demonstrate that the proposed AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.

ALMO DFT exploits non-orthogonal molecular orbitals (MOs) in a two-step self-consistent filed (SCF) approach that is designed to avoid the usual convergence issues.~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999}
We show that due to its linear scaling nature and relatively low computational prefactor, ALMO-based DFT calculations allow for simulation of very large weakly-coupled molecular systems. 
To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to balance force calculation errors.~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why} 
We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.

%\section{Theory}

Our approach is based on the ALMO SCF two-step variational method.~\cite{a:almo-ls} 
Initially, the electrons are partitioned into groups based on a logical approach, generally corresponding to molecules. 
In the first step, the basis atomic orbitals (AOs) are constrained to their parent molecules and a variational optimization is performed under the Born-Oppenheimer approximation using the DFT-SCF method within the non-overlapping domains. 
This is akin to calculating the electronic structure of each molecule (with frozen nuclei) within a coulombic field produced by the surrounding molecules. 
Charge transfer is completely constrained. 
In the second step, the variationally optimized but constrained MOs are allowed to delocalize onto nearby groups up to a desired cutoff radius $R_{c}$. 
This step requires a second variational optimization to be performed. 
Because of the logical partitioning of the groups and ignoring of long range forces, both steps of the electronic structure calculation are linear scaling with respect to system size. 
For more details on this method, see ref.\ \citenum{a:almo-ls}.

It is important to note that the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}. These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost. We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix:
%
\begin{eqnarray}
\label{eq:properdm}
\hat{R} = \sum_{x} \vert \psi^{xi} \rangle \langle \psi_{xi}\vert.
\end{eqnarray}
%
This approach provides the most rigorous description of the electronic structure.

A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations. 
Setting $R_{c}$ to a larger value reduces this discontinuity, but greatly increases the computational cost as longer distance (and less important) interactions must be calculated. 
As long as the discontinuity in force is small compared with the random error of the calculation, it can be safely ignored. Previously, this meant that $R_{c}$ had to be kept relatively large to maintain good accuracy of dynamical properties.

@@

\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{SCF}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}

Where $m_i$ refers to the mass of the i-th particle, and $r_{i\alpha}$ its position along dimension $\alpha$. 
$f^{\text{SCF}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma}_{i\alpha} (t)$. 
In order to correctly sample the canonical distribution, the random force must be zero mean. 
As well, the damping and white noise terms must be related by the second fluctuation dissipation theorem.~\cite{a:Kubo-1986,a:langevin-why,b:tuckerman-stat}
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This means the damping and white noise terms will work to maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities.

@@

The concept of balancing errors in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
Previously, $\epsilon_{SCF}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
Obviously, this introduces a significant error into the calculation of forces during each time step. 
We assume that this error has the form:
%
\begin{align}
\label{eq:assumption}
f^{\text{SCF}}_{i\alpha} = f^{\text{APP}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
\end{align}
%
where $f^{\text{SCF}}_{i\alpha}$ is the force calculated from an accurate and fully converged SCF method, $f^{\text{APP}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
Hence, the Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{APP}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
Note that $\Delta$ can be either positive or negative depending on the way errors are introduced in the system of interest. 
$\Delta$ was found to be positive for all systems considered here, increasing the amplitude of the effective random force on the system.

@@

\begin{figure}
\includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
\caption{\label{fig:forcecomp}The ``Random force'' $\langle R^{\Delta}_{i\alpha} \rangle (t)$ calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged across all dimensions and all molecules separately at each time step. 
Plotted are the random forces with varying localization radius $R_{c}$, in units of Van der Waals Radii.
Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} \rightarrow \infty$.
Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and the TZV2P basis set on 125 water molecules for 10 ps.}
\end{figure}

To test the accuracy of our unconverged ALMO based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory (see fig. \ref{fig:dynproperties} and supporting information). 
We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged KS DFT forces.
The kinetic energy distribution is compared with the theoretical Maxwell-Boltzmann distribution at the average temperature of the system.
From the slope of the MSD functions we also calculated diffusion constants.
Using fully converged KS DFT forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$, whereas with unconverged ALMO forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.

@@

%\textbf{Supporting Information}

>>>>

% Re-write the following section and include any recent advances
In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian, known as canonical KS orbitals, must be replaced with an alternative set of \emph{local} electronic descriptors
~\footnote{Canonical KS orbitals are delocalized over all atoms in the system. 
This delocalization is the main reason for the steep growth of the cost of the conventional KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom that describe delocalized orbitals grows quadratically.}. 
Most LS methods~\cite{a:linscale3,a:lee-yang-1996,a:ls-scuseria-1997,a:ls-manolopoulos-1998,a:ls-helgaker-2001,a:ls-niklasson-2003,a:curvy2,a:ls-dm-sign} explore the natural locality of the one-electron density matrix (DM). 
%However, the variational optimization of the DM is very inefficient for accurate DFT calculations that require many basis functions per atom~\cite{a:ls-rev-1999,a:ls-dm-sign}. 
However, in accurate DFT calculations that require many basis functions per atom, the variational optimization of the DM is advantageous only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}.
%However, in accurate DFT calculations with many basis functions per atom are required, the DM becomes sufficiently sparse only for impractically large systems~\cite{a:ls-rev-1999,a:ls-dm-sign,a:almo-ls}. 
Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems. 
The optimal basis variants of the DM methods~\cite{a:ls-stechel-1994,a:ls-gillan-1995,a:ls-gillan-1996,a:ls-onetep-2003} designed to rectify this issue contract large basis sets into a small number of new localized basis functions and then optimize the DM in the contracted basis. 
Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} their use in AIMD is hampered by the costly optimization of both the contracted orbitals and the density matrix. 
%
Alternatively, LS can be achieved using a set of local electronic descriptors constructed by enforcing strict locality on the KS orbitals and, at the same time, by relaxing the orthogonality constraints imposed on them. 
From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are more promising since they require only the occupied orbitals and, thus, deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the progress in the development of orbital-based LS methods has been hindered by the inherently difficult optimization of localized orbitals~\cite{a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:ls-rev-1999,a:weitao-yang-2013}.

%ZZZ-important addition: we did not include a large family of the so-called fragmentation methods 
%included only those methods that attempt to construct  set 
%It is important to note all LS DFT methods reviewed above attempt to obtain the electronic description that do not construct  based on fragmentation methods that do that there  LS DFT methods the presented logical partitioning with the subsequent construction of localized MOs is a rather general approach used in a large number of electronic structure theories, which are collectively known as fragmentation methods~\cite{a:fragmentation-rev}. These methods vary greatly in how the partitioning and recombination are performed and, therefore, differ in accuracy and computational cost. We would like to emphasize that in our approach a proper quantum mechanical description of the entire system is constructed in the form of the total idempotent density matrix. This approach provides the most rigorous description of the electronic structure.

%Thus, despite remarkable recent progress in linear scaling density function theory, the computational cost of existing methods remains too high for routine \emph{ab initio} molecular dynamics (AIMD) simulations.
Thus, despite remarkable recent progress the high computational overhead of existing asymptotically linear-scaling DFT methods restrict their use in AIMD simulations to very short \emph{time scales}, systems of low dimensions, or problems that can be adequately described with minimal basis sets. 
Otherwise, LS DFT cannot compete with the straightforward cubically scaling KS DFT on the length scales accessible in routine AIMD simulations.
%Thus,  the computational cost of the existing asymptotically linear-scaling DFT methods remains too high to compete with the straightforward cubically scaling KS DFT on the length scales practically accessible in routine AIMD simulations.
%Thus, although there exist numerous DFT methods capable of yielding asymptotic LS, their computational cost for systems in/on the practically accessible length scale remains too high to compete with the conventional cubically scaling KS DFT with its highly-optimized numerical algorithms.

In this work, we introduce a new AIMD method for weakly-interacting molecular systems that is not only linear scaling but also has a low computational overhead. 
Using liquid water as an example, we demonstrate that the proposed AIMD methodology enables computational studies of molecular systems on previously inaccessible time and length scales.

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own localization center -- atomic orbitals of a molecule are used here -- and a predefined localization radius that typically includes localization centers of neighbor molecules~\cite{a:stoll,a:almo-ls}. 
The key feature of ALMO DFT is that its trial wavefunctions are constructed to circumvent the problem of the sluggish variational optimization that have rendered many previous MO-based LS methods impractical. 
The trial ALMOs are optimized variationally in a two-step self-consistent-field (SCF) procedure~\cite{a:almo-ls}. 
In the first step, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second step, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second step, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first step. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

ALMO constraints restrict/prohibit/forbid electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes it, dispersion interactions~\cite{a:theeda}. 
Since the energetic contribution of the electron transfer decays exponentially with distance, \mbox{ALMO} approximation provides a natural and accurate representation of the electronic structure of molecular systems with just [a?] few local variables if an appropriate, that is sufficiently large, localization radius is chosen. 
%The advantage of ALMO DFT is that its computational complexity is linear with respect to system size because long-range electron delocalization is ignored in both optimization steps.
%Because of the greatly reduced number of electronic variables and their robust optimization the computational cost of the implemented ALMO DFT is so low that it is [has the potential] capable of extending the range of accurate simulations to systems of thousands of molecules.
Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT is linear with the number of molecules while its computational overhead remains very low. These features make ALMO DFT a promising method to perform accurate AIMD simulations of large molecular systems.

%To decrease the cost even further, the present paper builds on the ALMO SCF method as well as other recent work involving Brownian dynamics to compensate errors in the forces~\cite{a:ls-parinello,a:2ndcpmd,a:langevin-why}. 
%We will show that our approach lowers the computation cost pre-factor for the ALMO SCF method, while maintaining LS and accuracy for calculation of dynamical properties. 
%This is done by balancing a new computationally economical but systemically inaccurate force calculation method with an optimized version of the Langevin Equation of motion. 
%We also introduce an easy to follow heuristic approach to optimize a system for generating correct dynamical behaviour.

To adapt ALMO DFT for AIMD, the forces on atoms were computed \emph{approximately} using a straightforward procedure and the idempotent DM obtained from ALMOs (see the Supplemental Material). 
The difference between these forces and the perfect forces from perfectly-converged fully-delocalized KS orbitals is denoted $\delta f_{i\alpha}(t)$ 
%
\begin{align}
\label{eq:assumption}
f^{\text{KS}}_{i\alpha}(t) = f^{\text{ALMO}}_{i\alpha}(t) + \delta f_{i\alpha} (t),
\end{align}
%
where $\alpha$ is a Cartesian component of the force acting on atom $i$ at time step $t$. $\delta f_{i\alpha} (t)$ comprises neglected terms that originate from %fixed terms in the trial ALMOs (i.e. orbitals obtained in the first SCF step are not re-optimized in the second step), 
a finite localization radius $R_c$ and incomplete SCF optimization of the electronic degrees of freedom. $\delta f_{i\alpha} (t)$ can be reduced \emph{systematically} to zero by increasing $R_c$ (Figure~\ref{fig:forcecomp} in the Supplemental Material) and, at the same time, decreasing $\epsilon_{SCF}$ -- the maximum norm of the electronic energy gradient, below which the electronic ground state is considered sufficiently close to the Born-Oppenheimer state.

%A known source of error in force calculation within the ALMO SCF paradigm occurs due to the discontinuous jump in potential energy when a neighboring molecule passes beyond the localization radius $R_{c}$ and is suddenly ignored in potential energy calculations. 
Setting $R_{c}$ to a larger value reduces this discontinuity, but greatly increases the computational cost as longer distance (and less important) interactions must be calculated. 
As long as the discontinuity in force is small compared with the random error of the calculation, it can be safely ignored. Previously, this meant that $R_{c}$ had to be kept relatively large to maintain good accuracy of dynamical properties.

\begin{figure}
\includegraphics[trim={0.1cm 0cm 0.2cm 0.1cm},clip,width=8.6cm]{DeltaForceComparison_ALMO_SCF.eps}
\caption{\label{fig:forcecomp} {\color{red} All figures: VdWR$\rightarrow$vdWR. Hayden, your description states that you average over Cartesian components of forces. Unfortunately, averaging over components will give zero for any isotropic system observed for a sufficiently long time (i.e. for every component there will be equal but opposite component). What we would like to plot is the instantaneous magnitude of the vector force on atom/molecule, $\vert \delta \vec{f}_i(t) \vert$, averaged over all atoms/molecules.} Dependence of the instantaneous magnitude of $\vert \delta \vec{f}_i(t) \vert$ on the ALMO localization radius $R_c$. Forces are averaged over all atoms/molecules, $R_c$ is given in units of atomic van der Waals radii. %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step. Plotted are the random forces with varying localization radius $R_{c}$, .
%Here it is clear that ALMO DFT forces converge to Born-Oppenheimer forces in the limit of $R_{c} \rightarrow \infty$.
Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{SCF} =ZZZ$ for configurations from a 10-ps trajectory generated with KS DFT at T=298~K.}
\end{figure}

@@

\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}

Where $m_i$ refers to the mass of the i-th particle, and $r_{i\alpha}$ its position along dimension $\alpha$. 
$f^{\text{KS}}_{i\alpha}$ is the force calculated using the ALMO SCF method described above for the i-th particle along dimension $\alpha$, $\gamma$ is the Langevin scaling factor which sets the strength of both the damping term  $\gamma m_i \dot{r}_{i\alpha}$ and the stochastic Gaussian ``white noise'' term $R^{\gamma}_{i\alpha} (t)$. 
In order to correctly sample the canonical distribution, the random force must be zero mean. 
As well, the damping and white noise terms must be related by the second fluctuation dissipation theorem.~\cite{a:Kubo-1986,a:langevin-why,b:tuckerman-stat}
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This means the damping and white noise terms will work to maintain a constant temperature while also generating the correct Maxwell-Boltzmann distribution of velocities.

@@

The concept of balancing errors in force calculation with the Langevin equation can be taken even further, thereby reducing computational overhead even more. 
During the second step of the ALMO SCF method described above, a self-consistent variational approach is taken to converge the non-orthogonal MOs. 
Generally, this step utilizes a iterative search function in order to reduce the absolute norm of the energy gradient to some target value $\epsilon_{SCF}$. 
Previously, $\epsilon_{SCF}$ had to be set very small to ensure sufficient convergence of the electron structure calculations. 
By increasing the allowable $\epsilon_{SCF}$, a great reduction in computational cost can be achieved. 
Obviously, this introduces a significant error into the calculation of forces during each time step. 
We assume that this error has the form:
%
\begin{align}
\label{eq:assumption}
f^{\text{KS}}_{i\alpha} = f^{\text{ALMO}}_{i\alpha} + R^{\Delta}_{i\alpha} (t)
\end{align}
%
where $f^{\text{KS}}_{i\alpha}$ is the force calculated from an accurate and fully converged SCF method, $f^{\text{ALMO}}_{i\alpha}$ is the approximate force calculated by allowing a large $\epsilon_{SCF}$, and $R^{\Delta}_{i\alpha} (t)$ is a stochastic ``white noise'' term that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
It's clear from figure \ref{fig:randomforce} that this assumption is sufficiently justified. 
Hence, the Langevin equation of motion (\ref{eq:langevin}) can be re-written with the two stochastic terms combined into one:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha} - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
Note that $\Delta$ can be either positive or negative depending on the way errors are introduced in the system of interest. 
$\Delta$ was found to be positive for all systems considered here, increasing the amplitude of the effective random force on the system.

@@

To test the accuracy of our unconverged ALMO based MD we computed the Infrared (IR) spectrum, radial distribution function (RDF), kinetic energy distribution, and mean-squared displacement (MSD) function from a 10 ps MD trajectory (see fig. \ref{fig:dynproperties} and the Supplemental Material). 
We compared the IR spectrum, RDF, and MSD function to corresponding reference functions calculated with fully converged KS DFT forces.
The kinetic energy distribution is compared with the theoretical Maxwell-Boltzmann distribution at the average temperature of the system.
From the slope of the MSD functions we also calculated diffusion constants.
Using fully converged KS DFT forces we calculated D = $\SI{5.47e-9}{\m^{2}\per\s}$, whereas with unconverged ALMO forces we calculated D = $\SI{5.55E-09}{\m^{2}\per\s}$.

@@

%\textbf{Supplemental Material}

%ZZZ outline procedure to compute forces using the DM obtained from ALMOs.
% ZZZ figure to show
